### 2014-04-16 doc/uist14/iteration3.tex
iteration 3: more detailed description of orientation method
* Changed the structure slightly to discuss user facing first, then
  technical details.
* Discussed technical challenges in more detail
* Explained the "sticky" threshold more clearly (hopefully)
* Responded to some comments

%!TEX root = uist14.tex
\section{Iteration 3: Head motion refinement}
\label{sec:iteration-3:-head}
\subsection{Problem Statement and Goal}
Both{\em Naive IR} and {\em Intensity IR} rely on list navigation on the near-eye display for the refinement step. This has two problems: First, a spatial layout of target is compressed into a uni-dimensional list. List navigation commands thus have poor spatial mapping. Second, our techniques require the user to switch their focus back and forth between the physical world and the display. This issue motivated us to design an approach that also utilizes orientation information during the {\em refinement} stage, to provide a more direct mapping between the user's actions and target selection.

@@

\subsection{Technique}
The motion sensors (including gyroscope, magnetometer and accelerometer) on the head-worn device can help us determine the orientation of a user's head. However, this absolute orientation cannot be directly used in any indoor environment as the user may change their location. However, as long as the targets are spread around the periphery, their relative orientation doesn't change. This key observation inspires us to investigate a hybrid solution with IR and orientation sensors on Glass.

@@

In the first stage, the user orients their head towards each device, and the system saves the absolute orientations of each device. Once the user has added all devices, these orientations are saved for future use. It is still possible to add new devices at any point without running the entire calibration step, as the system will simply store the absolute orientation of the new device.

During the second stage, when the user wants to use head-motion for disambiguation, they will put their finger on the touchpad to enter a quasi-mode.  This gesture is inspired by observing users who tend to adjust their glasses for closer inspection.

In this mode, only one device has a solid indicator light on at a time. \bjoern{revisit prior sentence - there's a problem in the current technique in that the user doesn't know which targets are in the refinement set and which aren't.} \achal{There is no longer a refinement set - instead the user can select between all the devices using just orientation once they enter quasi mode.} The user can then tilt in the direction of the device they wish to switch to. We calculate the direction of movement using a low-pass-filtered history of orientations, and look for devices that lie in that direction. Once a new device is selected, the frame of reference changes, and the user can repeat the process. Note that in reality, this is a completely smooth operation: it is possible to switch between multiple devices in one continuous motion. This is achieved by automatically adjusting the frame of reference to the currently lit device, but also creating a ``sticky" threshold so that the user is not constantly jumping between devices. \bjoern{This is called hysteresis or double thresholding.}

\bjoern{how well will this work if the user shifts position within a room?}

>>>>

%!TEX root = uist14.tex
\section{Iteration 3: Head motion refinement}
\label{sec:iteration-3:-head}
\subsection{Problem Statement and Goal}
Both{\em Naive IR} and {\em Intensity IR} rely on list navigation on the near-eye display for the refinement step. This has two problems: First, a spatial layout of targets is compressed into a uni-dimensional list. List navigation commands thus have poor spatial mapping. Second, our techniques require the user to switch their focus back and forth between the physical world and the display. These issues motivated us to design an approach that also utilizes orientation information during the {\em refinement} stage, to provide a more direct mapping between the user's actions and target selection.

@@

\subsection{Technique}
The motion sensors (including gyroscope, magnetometer and accelerometer) on the head-worn device can help us determine the orientation of a user's head. However, this absolute orientation cannot be directly used in any indoor environment as the user may change their location. However, as long as the targets are spread around the periphery, their relative orientation is stable. This key observation inspires us to investigate a hybrid solution with IR and orientation sensors on Glass.

@@

In the first stage, the user orients their head towards each device, and the system learns a map of the devices based on their corresponding sensor readings. During the second stage, when the user wishes to use head-motion for disambiguation, they will put their finger on the touchpad to enter a quasi-mode in which their head movement is used to select targets. In this mode, only one device has a solid indicator light on at a time. \bjoern{revisit prior sentence - there's a problem in the current technique in that the user doesn't know which targets are in the refinement set and which aren't.} \achal{There is no longer a refinement set - instead the user can select between all the devices using just orientation once they enter quasi mode.} The user can then tilt in the direction of the device they wish to switch to. This gesture is inspired by observing users who tend to adjust their glasses for closer inspection.

The technical implementation of this interaction involves a number of subtleties due to the noisiness of our data. Absolute orientations between devices change frequently at the slightest user movement, and motion sensor data is inherently noisy. To alleviate the second issue, a low-pass filter is applied to the measurements, allowing us to extract the direction of motion from this smoothed history. Empirically, we found that we can reliably detect the direction of movement with 100ms of history and a sensor resolution of 10ms.

Orientation changes can be overcome by using a hybrid of absolute and relative orientations augmented by IR information. While absolute orientations are unstable, relative orientations do not contain information for essential tasks, such as adding devices to a learned map. Our system stores the absolute orientations of each device during the learning stage, but utilizes only the relative information embedded in this data for disambiguation. Once the user is in quasi-mode, our system calculates the user's direction of motion as described earlier, and searches through the map for the nearest device in that direction. Thus, although the absolute orientation of the currently lit device may be incorrect, the relative information can still be used for disambiguation. Finally, the system uses hysteresis to prevent the user from accidentally jumping repeatedly between devices by switching devices only once the user has moved beyond a certain threshold in any direction. \achal{I understand hysteresis, but I can't figure out how the word can be used naturally (as there's no way to say a system "has" hysteresis). This last sentence may require rewording.}

\bjoern{how well will this work if the user shifts position within a room?}
\achal{I think the updates should explain this now.}
